Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-07 paper/paper_new.tex
Edited Section 5.

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 100 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

Our automated process currently generates 6 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), and \verb|brightness| (\verb|foreground.mean|).
% and \verb|if_background_uniform| (\verb|background.std|).
In addition, our process also calculates contrast, bounding boxes and
fish lengths in centimeters.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\linewidth]{images/none1_crop} \\
  \vspace{3mm}
  \includegraphics[width=0.9\linewidth]{images/double1_crop}
  \caption{A fish that was not detected (top) and a fish that was detected twice (bottom).}
  \label{fig:fish_detect_wrong}
\end{figure}

@@

Images where one of these objects were not detected generally had some form of coloration issue. They were either washed out, very dark or yellow in hue.
See Figure \ref{fig:no_ruler} two such examples.
Some of the rulers for which ``2'' and/or ``3'' were not detected were particularly scratched and damaged. Only ``3'' was missed in 45 of the 56 cases, only ``2'' was missed in 2 cases, and both were missed in 9 cases. This may indicate that more training samples for ``3'' are required. Many of the rulers where both numerals were missed were particularly small within the image, which again may be solvable through expanding the training dataset.

@@

% There were six cases where an eye was detected but the \verb|side| value
% was wrong.
For these 6 cases, a spot on the wrong side of the fish was labeled as the
most likely eye within the bounding box of the fish, i.e.\ eye detection 
was incorrect. Figure \ref{fig:wrong_eye} presents one such example.
There were an additional 17 images for which the automated process generated a result that did not match the manually created data. For these remaining cases the manual data was incorrect, giving the automated system an error rate \(2.8\) times lower than the human error rate.
This result highlights the additional utility of automated methods to
double-check and verify human-generated metadata.

@@

Of the specimens for which clock values were generated, 33 did not match the manually created data (within a tolerance of \(\pm{}1\)).
For 25 of those, the human-generated data was incorrect, giving the automated process a \(3.1\) times lower error rate. Of the 8 that were
computationally classified incorrectly,
two specimens were quite curved making it difficult to assign a clear angle value, as
seen in Figure \ref{fig:curvedFish}.
The other 6 were values of ``3'' instead of ``9'' or vice versa, which resulted from
a mislabeled eye as discussed in the previous section.
%\todo{Why is the following sentence here? What does it have to do with
%Clock Value?}
%A small number of images (such as on the right in Figure~\ref{fig:teaser}) have tags that overlap with the fish. These tags were light enough that they were correctly labeled as background by the pixel analysis.

@@

\subsection{Image Contrast}
The contrast of an image is an important image property needed
for the analyses of the BGNN study. Ideally, for INHS images the fish should be well lit and clearly displayed, and the background should also be as light and white as possible.
% Figure~\ref{fig:histo} shows the distribution of
The difference between
the mean intensity of the foreground (i.e.\ the pixels of the fish) and the mean intensity of the background within the fish's bounding box
was computed for all 7,244 images.
The overall mean of the differences is \(144.3\), with a standard deviation of \(15.8\). Images on the low end of the distribution exhibit poor foreground--background contrast, and images on the high end likely contain poorly lit specimens.
Images are considered to have ``low'' and
``high'' contrast if their background-foreground difference is greater than
one standard deviation away from the mean; otherwise they are classified as
``medium''.
Examples of each type of image (low, medium, and high contrast) can be seen in Figure~\ref{fig:contrast}.
The left image has a mean intensity of 103 (low), the middle image 144
(medium), and the right 186 (high).

\begin{table}[H]
    \centering
%    \caption{\verb|foreground.mean| statistics for the three Brightness classes}
    \caption{foreground.mean statistics for the three Brightness classes}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.2 & 13.5 & 1800\\
        Normal & 93.6 & 14.8 & 5186\\
        Bright & 108.4 & 15.2 & 242\\
        \bottomrule
    \end{tabular}
\end{table}
\subsection{Brightness}
Specimen brightness is one of the 22 hand-recorded metadata properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values
correspond to the mean foreground intensity computed by the automated system.
The mean and standard deviation of the foreground intensities were
computed for the images in the three human-specified classes.
Table~\ref{tab:bright} contains the resulting values and shows that
automated intensities values provide objective measures that may be used to break images into groups that correlate with human-defined brightness
classifications.
The mean of \verb|foreground.mean| over all 7,244 images is 89.5, with a
standard deviation of 16.6. These numbers suggest, for example, that
\verb|foreground.mean| $\pm$ one-half a standard deviation could define 
\verb|normal| brightness, with values outside this range being classified
as \verb|dark| or \verb|bright|.

@@

Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with those 263 falling back to the original \verb|detectron| bounding box.
100 randomly-chosen images were reviewed manually to evaluate the
calculation, a representative sample are presented in
Figure \ref{fig:bbox_mask}.
All 100 masks and bounding boxes were correctly placed
on/around the location of the fish. However, a number of them lacked portions of lightly colored tails and/or fins.
Specifically, 41 masks and bounding boxes covered the entire fish,
36 missed some of the tail, and 23 missed most or all of the tail, as
seen in Figure \ref{fig:missedTail}.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/61631}
  \includegraphics[width=0.49\linewidth]{images/61631_mask}
  \caption{An example of a light colored tail being missed during the pixel analysis process.}
  \label{fig:missedTail}
\end{figure}
Masks and bounding boxes contain the head and trunk of the fish in nearly all cases, but further refinement of our algorithms will be needed
to ensure that light fins and tails are masked consistently and accurately.

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images.
For the remaining 65 images, either the fish, the ``2'' and/or the ``3'' on the ruler were not detected.
Image scale ($\frac{\mathrm{pixels}}{\mathrm{cm}}$) and fish length were
measured, using ImageJ~\cite{imagejCite}, in the same 100  test images.
% 50 of these images \todo{The same 50 as before?} were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ~\cite{imagejCite}.
In this subset of images the average error for the scale calculation
was \(0.91\%\), and the average error for the fish length calculation was \(5.87\%\).
Scale calculations using the ``2'' and ``3'' method are nearly identical to those calculated by hand between the tick marks on the ruler. When the tail of the fish is accurately masked and the specimen is fairly straight, the length calculation is highly accurate as well. An example of such a result can be seen in Figure~\ref{fig:scale_len}, for which the difference between the hand measured length and the automatically calculated length was only \(0.6\%\) (\(8.88\) cm vs \(8.82\) cm). Thus, the primary means of lowering the error of the length calculation is to improve tail masking accuracy.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/54172}
  \includegraphics[width=0.49\linewidth]{images/54172_mask}
  \caption{A fish for which the masking and measurement process was highly accurate.}
  \label{fig:scale_len}
\end{figure}

>>>>

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction along its major
axis. Since fish are normally measured in a straight line from their snout down the middle of their trunk, every pixel of the fish blob is projected
onto the fish's major axis (as a line through its centroid).
The projection is done by finding the closest point on the
centroid--principal axis line from the pixel's location.
% intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line.
After processing every pixel in the fish blob, computing the distance
between the two furthest projected points gives the length of the fish in pixels. Multiplying this distance by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}

\begin{figure}[b]
  \centering
  \includegraphics[width=0.9\linewidth]{images/none1_crop} \\
  \vspace{3mm}
  \includegraphics[width=0.9\linewidth]{images/double1_crop}
  \caption{A fish that was not detected (top) and a fish that was detected twice (bottom).}
  \label{fig:fish_detect_wrong}
\end{figure}

\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{leipzig2021biodiversity} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,244\) were both not part of the \verb|detectron| training set and met our current admissibility criteria for \verb|detectron| and pixel processing. We ran the metadata extraction program on these \(7,244\) images.
For the properties of image scale, fish length, and fish bounding boxes
(properties not manually generated),
a random sample of 100 specimens from the set of \(7,244\) were analyzed by
hand for comparison.

Our automated process currently generates 6 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), and \verb|brightness| (\verb|foreground.mean|).
% and \verb|if_background_uniform| (\verb|background.std|).
In addition, our approach also calculates contrast, bounding boxes and
fish lengths in centimeters.

@@

Images where one of these objects were not detected generally had some form of coloration issue. They were either washed out, very dark or yellow in hue.
See Figure \ref{fig:no_ruler} for two such examples.
Some of the rulers for which ``2'' and/or ``3'' were not detected were particularly scratched and damaged. Only ``3'' was missed in 45 of the 56 cases, only ``2'' was missed in 2 cases, and both were missed in 9 cases. This may indicate that more training samples for ``3'' are required. Many of the rulers where both numerals were missed were particularly small within the image, which again may be solvable through expanding the training dataset.

@@

% There were six cases where an eye was detected but the \verb|side| value
% was wrong.
For these 6 incorrect cases, a spot on the wrong side of the fish was
labeled as the most likely eye within the bounding box of the fish.
% , i.e.\ eye detection was incorrect.
Figure \ref{fig:wrong_eye} presents one such example.
There were an additional 17 images for which the automated process generated a result that did not match the manually created data. For these remaining cases the manual data was incorrect, giving the automated system an error rate \(2.8\) times lower than the human error rate.
This result highlights the additional utility of automated methods to
double-check and verify manually generated metadata.

@@

Of the specimens for which clock values were generated, 33 did not match the manually created data (within a tolerance of \(\pm{}1\)).
For 25 of those, the manually generated data was incorrect, giving the automated process a \(3.1\) times lower error rate. Of the 8 that were
computationally classified incorrectly,
two specimens were quite curved making it difficult to assign a clear angle value, as
seen in Figure \ref{fig:curvedFish}.
The other 6 were values of ``3'' instead of ``9'' or vice versa, which resulted from
a mislabeled eye as discussed in the previous section.
%\todo{Why is the following sentence here? What does it have to do with
%Clock Value?}
%A small number of images (such as on the right in Figure~\ref{fig:teaser}) have tags that overlap with the fish. These tags were light enough that they were correctly labeled as background by the pixel analysis.

@@

\subsection{Image Contrast}
The contrast of an image is an important image property needed
for the analyses of the BGNN study. Ideally, for INHS images the fish should be well lit and clearly displayed, and the background should also be as light and white as possible.
% Figure~\ref{fig:histo} shows the distribution of
The difference between
the mean intensity of the foreground (i.e.\ the pixels of the fish) and the mean intensity of the background within the fish's bounding box
was computed for all 7,244 images.
The overall mean of the differences is \(144.3\), with a standard deviation of \(15.8\). Images on the low end of the distribution exhibit poor foreground--background contrast, and images on the high end likely contain poorly lit specimens.
Images are considered to have ``low'' and
``high'' contrast if their background-foreground difference is greater than
one standard deviation away from the mean; otherwise they are classified as
``medium''.
Examples of each type of image (low, medium, and high contrast) can be seen in Figure~\ref{fig:contrast}.
The left image has a contrast of 103 (low), the middle image 144
(medium), and the right 186 (high).

\begin{table}[H]
    \centering
%    \caption{\verb|foreground.mean| statistics for the three Brightness classes}
    \caption{foreground.mean statistics for the three Brightness classes}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.2 & 13.5 & 1800\\
        Normal & 93.6 & 14.8 & 5186\\
        Bright & 108.4 & 15.2 & 242\\
        \bottomrule
    \end{tabular}
\end{table}
\subsection{Brightness}
Specimen brightness is one of the 22 hand-recorded metadata properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values
correspond to the mean foreground intensity computed by the automated system.
The mean and standard deviation of the foreground intensities were
computed for the images in the three manually specified classes.
Table~\ref{tab:bright} contains the resulting values and shows that
automated intensities values provide objective measures that may be used to break images into groups that correlate with manually generated
brightness classifications.
The mean of \verb|foreground.mean| over all 7,244 images is 89.5, with a
standard deviation of 16.6.
Given that human perception of brightness is quite subjective and
highly variable, we found it problematic to use specific threshold
values to define \verb|dark|, \verb|normal| or \verb|bright| specimens
in concordance with the manually generated metadata.  The technicians
showed little consistency when classifying \verb|normal| or \verb|bright|
specimens. We did find that an 81.2\% accuracy rate could be achieved by
classifying \verb|dark|, as those with a \verb|foreground.mean| of 75
or lower; thus demonstrating that this computed metadata property does 
offer some value when assessing image quality.
% These numbers suggest, for example, that
% \verb|foreground.mean| $\pm$ one-half a standard deviation could define 
% \verb|normal| brightness, with values outside this range being classified
% as \verb|dark| or \verb|bright|.

@@

Fish bounding boxes were calculated for all \(7,237\) images in which a fish was found. All but 263 of these were generated via pixel analysis, with
the 263 falling back to the original \verb|detectron| bounding box.
100 randomly-chosen images were reviewed manually to evaluate the
calculation, a representative sample are presented in
Figure \ref{fig:bbox_mask}.
All 100 masks and bounding boxes were correctly placed
on/around the location of the fish. However, a number of them lacked portions of lightly colored tails and/or fins.
Specifically, 41 masks and bounding boxes covered the entire fish,
36 missed some of the tail, and 23 missed most or all of the tail, as
seen in Figure \ref{fig:missedTail}.
Masks and bounding boxes contain the head and trunk of the fish in nearly all cases, but further refinement of our algorithms will be needed
to ensure that light fins and tails are masked consistently and accurately.
\begin{figure}[t]
  \centering
  \includegraphics[width=0.49\linewidth]{images/61631}
  \includegraphics[width=0.49\linewidth]{images/61631_mask}
  \caption{An example of a light colored tail being missed during the pixel analysis process.}
  \label{fig:missedTail}
\end{figure}

\subsection{Scale and Length}
Image scale and fish lengths were calculated for \(7,179\) of the images.
For the remaining 65 images, either the fish, the ``2'' and/or the ``3'' on the ruler were not detected.
Image scale ($\frac{\mathrm{pixels}}{\mathrm{cm}}$) and fish length were
measured, using ImageJ~\cite{imagejCite}, in the same 100  test images.
% 50 of these images \todo{The same 50 as before?} were measured by hand to get \(\frac{\mathrm{pixels}}{\mathrm{cm}}\) and fish length in centimeters with ImageJ~\cite{imagejCite}.
In this subset of images the average error for the scale calculation
was \(0.91\%\), and the average error for the fish length calculation was \(5.87\%\).
Scale calculations using the ``2'' and ``3'' method are nearly identical to those calculated by hand between the tick marks on the ruler. When the tail of the fish is accurately masked and the specimen is fairly straight, the length calculation is highly accurate as well. An example of such a result can be seen in Figure~\ref{fig:scale_len}, for which the difference between the hand-measured length and the automatically calculated length was only \(0.6\%\) (\(8.88\) cm vs \(8.82\) cm). Thus, the primary means of lowering the error of the length calculation is to improve tail masking accuracy.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.49\linewidth]{images/54172}
  \includegraphics[width=0.49\linewidth]{images/54172_mask}
  \caption{A fish for which the masking and measurement process was highly accurate.}
  \label{fig:scale_len}
\end{figure}
